Open-source landscape for 3D CSEM modelling

### 2020-09-21 manuscript/Manuscript.tex
Final SimPEG block-model results
Updates from Overleaf

Furthermore, we deal with the topic of verification. Analytical and semi-analytical solutions only exist for simple halfspace or layered-Earth models, which served mostly to verify new codes. Beyond these simple models, the only objective possibility to ensure the accuracy of solutions is by comparing results from different modellers. If different discretizations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy. \cite{GJI.13.Miensopust} presents a review of two workshops dealing with the verification of magnetotelluric forward and inversion codes, but we are not aware of any comparable study or benchmark suite for CSEM data. This is, to the best of our knowledge, the first comparison exercise for CSEM codes. We hope our efforts help to ensure the reliability of simulations for any new or existing, open-source or closed-source 3D CSEM project.

@@

\section{Numerical Verification}

We computed the responses for three different models to verify that the four 3D CSEM codes yield the same electromagnetic responses and to compare different solver types (FE, FD) and different mesh types (unstructured tetrahedra, tensor, octree) in different scenarios. The first model is an anisotropic layered model, which can be compared to semi-analytical solutions. The layered model serves also as background model for the second verification, where we add three resistive blocks into the subsurface. The final comparison is based on the realistic, anisotropic marine model MR3D, which is an open resistivity model which comes with computed CSEM responses.

@@

The results for the three receiver lines $y=-3,0,+3\,$km are shown in the left, middle, and right columns of Figure~\ref{fig:results-block}, respectively. The top row shows the result of \emg3d, as an example, and the bottom row shows the NRMSDs between the absolute responses of the codes.
%
\plot*{results-block}{width=\textwidth}{Results of the block model comparison: The responses of \emg3d, as an example, are shown in the top row, and the NRMSDs (\%) between the absolute responses ($|E_x|$) of the different codes are shown in the bottom row.}
%
\itodo{We might wait for the final \simpeg results before jumping to conclusions; but generally, the NRMSD differences is below 1-2\,\%, except for large offsets.}

The corresponding required runtime and memory are listed in Table~\ref{tbl:comp-block}.
%
\tabl[btp]{comp-block}{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the block model.}{
  \centering
\begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
\toprule
  %
       & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  %
  \custem & 24 &   158 & 115.9 & 1654242 \\ % dof: 3308484 / 2  (Real+Imag -> Complex)
  \emg3d  &  1 &   115 &   0.4 & 4813216 \\
  \petgem & 24 &    72 &  35.8 & 2455868 \\
  \simpeg &  4 &  9993 & 280.6 & 4813216 \\ % nprocs: 1*4 threads
  %
  \bottomrule
\end{tabular}}%
%

@@

The Marlim oil field is a giant reservoir in a turbidite sandstone horizon in the north-eastern part of the Campos Basin, offshore Brazil, which was discovered in 1985. \cite{BJG.17.Carvalho} created from seismic data and well log data a realistic, three-dimensional resistivity model with vertical transverse isotropy (VTI), called MR3D, which they released under the open creative common license \emph{CC BY 4.0}. \cite{GEO.19.Correa} computed CSEM data for MR3D for six frequencies from 0.125\,Hz to 1.25\,Hz, and released them under the same CC license. To compute the data they used a state-of-the-art code from the industry \citep[][ \emph{SBLwiz} software from \emph{EMGS}]{GEO.07.Maao}. It is therefore an ideal case to verify our open-source codes against, as it is a complex, realistic model and the data were computed by an industry-proofed code. Additionally, that code is a time-domain code, whereas the four codes under consideration here compute the results all in the frequency-domain.

@@

Even if modelling results may look completely valid, they could not be correct at all. Verifying the performance of 3D CSEM codes for real 3D problems is only possible by cross-validating multiple solutions. Overall, we observed an excellent match between all solutions. The relative misfits, almost always smaller than 10\,\%, were related to the either weak or strong amplitudes of real or imaginary parts and can be attributed to particularities regarding the discretization, boundary effects, and interpolation. Considering this satisfactory result in terms of accuracy and robustness, it is more interesting to discuss the characteristics of the four considered codes with focus on the demand of computational resources and the capability of handling the investigated CSEM geometries, including the very important underlying mesh design. The following arguments are based on a comparison of only marine CSEM models, which are the focus of industry applications. Nevertheless, we are confident that many observations would hold in a similar manner for land-based, airborne, or mixed setups.

@@

@@

We compared four different open-source 3D CSEM modellers by computing responses for a layered, anisotropic model, including a modification with three resistive blocks, and the realistic marine Marlim R3D model. All comparisons exhibited an excellent reliability of the solutions. Our data should make it very easy for new codes to have a readily available data set to test against and verify. We are confident that the discussions about runtimes, memory consumption, solver choices, and, most important, discretizations of the 3D CSEM problems help potential users not only to decide which modeller is best suited for specific tasks, but also to support their individual projects.

>>>>

Furthermore, we deal with the topic of validation. Analytical and semi-analytical solutions only exist for simple halfspace or layered-Earth models, which served mostly to verify new codes. Beyond these simple models, the only objective possibility to ensure the accuracy of solutions is by comparing results from different modellers. If different discretizations and implementations of Maxwell's equations yield the same result, it gives confidence in their accuracy. \cite{GJI.13.Miensopust} presents a review of two workshops dealing with the validation of magnetotelluric forward and inversion codes, but we are not aware of any comparable study or benchmark suite for CSEM data. This is, to the best of our knowledge, the first comparison exercise for CSEM codes. We hope our efforts help to ensure the reliability of simulations for any new or existing, open-source or closed-source 3D CSEM project.

@@

\section{Numerical Validation}

We computed the responses for three different models to validate that the four 3D CSEM codes yield the same electromagnetic responses and to compare different solver types (FE, FD) and different mesh types (unstructured tetrahedra, tensor, octree) in different scenarios. The first model is an anisotropic layered model, which can be compared to semi-analytical solutions. The layered model serves also as background model for the second validation, where we add three resistive blocks into the subsurface. The final comparison is based on the realistic, anisotropic marine model MR3D, which is an open resistivity model which comes with computed CSEM responses.

@@

The results for the three receiver lines $y=-3,0,+3\,$km are shown in the left, middle, and right columns of Figure~\ref{fig:results-block}, respectively. The top row shows the result of \emg3d, as an example, and the bottom row shows the NRMSDs between the absolute responses of the codes.
%
\plot*{results-block}{width=\textwidth}{Results of the block model comparison: The responses of \emg3d, as an example, are shown in the top row, and the NRMSDs (\%) between the absolute responses ($|E_x|$) of the different codes are shown in the bottom row.}
%
The NRMSD is generally below a few percent, except for large offsets. An interesting feature is that the NRMSD between \emg3d and \simpeg is very low except at the boundary. The reason is that the codes use the same tensor mesh for modelling, but while \emg3d uses the PEC boundary condition, while \simpeg uses the PMC boundary condition.
\mycom[DW]{Lindsey, what about the different formulations we use?}
\todo{expand the analysis of the Block Model -- any ideas / inputs?}

The corresponding required runtime and memory are listed in Table~\ref{tbl:comp-block}.
%
\tabl[btp]{comp-block}{Comparison of number of processes, runtime, and memory, as well as the degree of freedom of the discretization used by the different codes for the block model.}{
  \centering
\begin{tabular}{lrS[table-format=6.0]S[table-format=4.1]S[table-format=8.0]}
\toprule
  %
       & \#Procs & {Runtime (s)} & {Memory (GiB)}   & {\#dof} \\
  \midrule
  %
  \custem & 24 &   158 & 115.9 & 1654242 \\ % dof: 3308484 / 2  (Real+Imag -> Complex)
  \emg3d  &  1 &   115 &   0.4 & 4813216 \\
  \petgem & 24 &    72 &  35.8 & 2455868 \\
  \simpeg &  4 & 10295 & 280.4 & 4813216 \\ % nprocs: 1*4 threads
  %
  \bottomrule
\end{tabular}}%
%

@@

The Marlim oil field is a giant reservoir in a turbidite sandstone horizon in the north-eastern part of the Campos Basin, offshore Brazil, which was discovered in 1985. \cite{BJG.17.Carvalho} created from seismic data and well log data a realistic, three-dimensional resistivity model with vertical transverse isotropy (VTI), called MR3D, which they released under the open creative common license \emph{CC BY 4.0}. \cite{GEO.19.Correa} computed CSEM data for MR3D for six frequencies from 0.125\,Hz to 1.25\,Hz, and released them under the same CC license. To compute the data they used a state-of-the-art code from the industry \citep[][ \emph{SBLwiz} software from \emph{EMGS}]{GEO.07.Maao}. It is therefore an ideal case to validate our open-source codes against, as it is a complex, realistic model and the data were computed by an industry-proofed code. Additionally, that code is a time-domain code, whereas the four codes under consideration here compute the results all in the frequency-domain.

@@

Even if modelling results may look completely valid, they could not be correct at all. Validating the performance of 3D CSEM codes for real 3D problems is only possible by cross-validating multiple solutions. Overall, we observed an excellent match between all solutions. The relative misfits, almost always smaller than 10\,\%, were related to the either weak or strong amplitudes of real or imaginary parts and can be attributed to particularities regarding the discretization, boundary effects, and interpolation. Considering this satisfactory result in terms of accuracy and robustness, it is more interesting to discuss the characteristics of the four considered codes with focus on the demand of computational resources and the capability of handling the investigated CSEM geometries, including the very important underlying mesh design. The following arguments are based on a comparison of only marine CSEM models, which are the focus of industry applications. Nevertheless, we are confident that many observations would hold in a similar manner for land-based, airborne, or mixed setups.

@@

\itodo{Discussion poins from Evert: How can you attribute differences to the solver? There are mesh element effects as well and the way Maxwell's equations are discretized plays a role too.}

@@

We compared four different open-source 3D CSEM modellers by computing responses for a layered, anisotropic model, including a modification with three resistive blocks, and the realistic marine Marlim R3D model. All comparisons exhibited an excellent reliability of the solutions. Our data should make it very easy for new codes to have a readily available data set to test against and validate. We are confident that the discussions about runtimes, memory consumption, solver choices, and, most important, discretizations of the 3D CSEM problems help potential users not only to decide which modeller is best suited for specific tasks, but also to support their individual projects.
